\documentclass[a4paper]{master}

\addbibresource{main.bib}

%Information as title and author
\title{Simulation of Jet D'Eau from A to Y:\\ Numerical algorithm to solve incompressible Navier-Stokes equations
with \textcolor{blue}{m}arker \textcolor{blue}{a}nd \textcolor{blue}{c}ell based method}
\author{Pablo Strasser}
\Supervisor{Prof. Martin Gander and Dr. Felix Kwok}
\date{\today}

%Pdf information
\hypersetup{pdfinfo={%
  Title={Master Thesis},
  Author={Pablo Strasser},
  Creator={Pablo Strasser},
  Producer={Pablo Strasser},
  Subject={Navier-Stockes},
  Keywords={Navier-Stockes}}
}

\tikzexternalize

\begin{document}
\captionsetup{singlelinecheck=off,margin=10pt,font=small,labelfont=bf}
%make the title page
\maketitle
\dominitoc

\chapter*{Abstract}

\mtcaddchapter[Abstract]

We will in this work present how to solve the Navier-Stokes equations with a marker and cell method.
We use direct the Navier-Stokes approach without other models than the Navier-Stokes equations.
We emphasize the mathematical modularity which allows to work separately on every piece of this work.
Our method works for every Runge-Kutta method (explicit or implicit). If we denote by $k$ the order of the chosen Runge-Kutta method,
our method is of order $k$ if we ignore boundary conditions, of order $k$ in the better case and of order $1$ for specific complicated case.
Our method is exactly divergence free (up to roundoff error).

We will specifically be interested in free surface boundary conditions. The final goal is to simulate a Jet d'Eau.
This goal was only partially achieved, but the needed basics are there. 


\tableofcontents

\chapter*{Introduction}

\mtcaddchapter[Introduction]

This work consists of the following chapters:
\begin{enumerate}
 \item A notation and basic math property chapter.
 \item A chapter that describes the problem analytically. In this chapter no discretization is done, but equations are written in the good form to motivate discretization.
 \item A chapter that focuses on the solution of Navier-Stokes equations if the domain of computation is fixed.
 \item A chapter that indicates the change needed to allow domain change.
 \item To end, a conclusion chapter which will contain hint and discussion to how to upgrade the scheme used.
\end{enumerate}

The two numerical chapters contain pseudo code that implement the described algorithm.

The original \textbf{C++11} (c++ code respecting the 2011 norm ISO/IEC 14882:2011) code can be found on \url{https://github.com/Strasser-Pablo/Mac}
but is not needed to understand this work, the pseudo code is sufficient.

All the chapters are self contained,
but further information and inspiration come from:
\begin{itemize}
 \item Algorithmic ideas like the layer variable and the first scheme come from \cite{fluidforrestofus}.
 \item Analytical comprehension and better formulation come from \cite{citeulike:11163809}.
 \item Extrapolation for boundary conditions come from \cite{FLD:FLD148}.
 \item Analytical solutions are taken from \cite{Batchelor} and \cite{Kampanis:2006:SGH:1148052.1148065}.
 \item A review of marker and cell based methods can be found in \cite{citeulike:3055787}.
\end{itemize}



\include{introduction/main}
\include{analytical/main}
\include{fixed/main}
\include{topology/main}
\include{Conclusion/main}
\printbibliography[heading=bibintoc]
\end{document}