\chapter{Analytical problem and analytical result}

\minitoc
\section*{Introduction}
In this chapter, we will discuss the analytical problem that we want to solve.
The other numerical chapters will consist of numerical schemes to solve the problem given in this chapter.

\section{Variables and parameters}

The problem is to find a function of space and time in a certain domain respecting some conditions and equations.

We begin to define space, time and domain:
\begin{description}
\item[The \emph{space}] consists of a vector in fixed Cartesian coordinates with dimension typically 2 or 3.
Space is denoted by $\vect{x}$.
\item[The \emph{time}] consists of a parameter indicating the evolution of the function in time.
Time is denoted by $t$.
\item[The \emph{domain}] is the spatial region where a field is defined. This consists physically in the spatial region where water lies.
Boundary conditions need to be provided.
The domain can vary with respect to time and can vary with respect to the solution at previous times.
The domain does not need to be connected, but can be decomposed into a union of connected subdomains.

\begin{example}
 Take a drop of water falling into a glass of water, the domain consists of the region where the water lies.
 At the beginning we have a domain consisting of two disconnected subdomains (the water in the drop and the water in the glass)
 that after a given time will merge into one domain.
\end{example}

The domain is generally denoted by $\Omega$ or $\Omega(t)$ to emphasize that it can be variable with respect to time.
\end{description}

The variable of the equation is named \emph{Field} and consists of functions with respect to space in the spatial domain and time in the temporal domain,
\begin{equation}
 F(x,t)\qquad \text{with $x\in\Omega(t)$ and $t_0\leq t \leq t_f$.}
\end{equation}

We have the following fields:
\begin{description}
\item[The \emph{speed}:] A vector field consisting of the speed in space and time.
The speed is denoted by $\vect{v}$.
\item[The \emph{pressure}:] A scalar field consisting of an internal force per unit area in the fluid.
We will not use the pressure in this work, but the pressure divided by density.
If we denote the pressure by $\tilde{p}$ we use
\begin{equation}
 p=\frac{\tilde{p}}{\rho}.
\end{equation}
Because we use the incompressible Navier-Stokes equation, $\rho$ is constant. This can be viewed as a change of units.
\end{description}

The following constant parameters are used:
\begin{description}
\item[The \emph{viscosity}] represents how much water particles stick together.
The higher the viscosity, the less abrupt speed changes are possible.
The viscosity is denoted by $\mu$. We use $\nu$ that is the viscosity divided by the density.
\begin{equation}
 \nu=\frac{\mu}{\rho}.
\end{equation}
\item[The \emph{density}] is a mass per unit volume.
Density plays the same role as mass in Newtonian mechanics.
In the incompressible Navier-Stokes equation density is a constant because incompressibility tells us that density does not change.
The density is denoted by $\rho$.
\item[The \emph{external force}] is a user defined vector field that represents the external force applied to the water.
It can possibly depend on the solution at previous times.
The force is denoted by $\vect{F}$.
\end{description}

\section{Navier-Stokes equations}

\subsection{Fixed Eulerian topology}
\label{analytical:fixe_eulerian}
The incompressible Navier-Stokes equations are given by
\begin{subequations}
\begin{align}
\label{analytical:navierStokesA}
\vect{\nabla} \cdot \vect{v}(\vect{x} ,t)&=0,\\
\label{analytical:navierStokesB}
\partial_t \vect{v}(\vect{x} ,t)+(\vect{v}(\vect{x},t)\cdot\vect{\nabla} ) \vect{v}(\vect{x} ,t)&=-\vect{\nabla} p(\vect{x},t)+\frac{\vect{F}(\vect{x},t)}{\rho(\vect{x},t)}+\nu \Delta \vect{v}(\vect{x},t).
\end{align}
\end{subequations}
Equation (\ref{analytical:navierStokesA}), tells us that the speed is a divergence free field.
Equation (\ref{analytical:navierStokesB}) is an equation of evolution. The presence of $p$ allows to maintain the speed always divergence free.
Taking the curl of (\ref{analytical:navierStokesB}) makes disappear the pressure term (because the curl of a gradient is 0).

The pressure term can be seen as a correction of the divergence without changing the curl.

We can rewrite the Navier-Stokes equations as
\begin{subequations}
\begin{align}
\label{analytical:navierStokesRewA}
\vect{\nabla} \cdot \vect{v}(\vect{x} ,t)&=0,\\
\label{analytical:navierStokesRewB}
\partial_t \vect{v}(\vect{x} ,t)&=f(\vect{v}(\vect{x},t))-\vect{\nabla}p,
\intertext{where}
f(\vect{v}(\vect{x},t))&=-(\vect{v}(\vect{x},t)\cdot\vect{\nabla} ) \vect{v}(\vect{x} ,t)+\frac{\vect{F}(\vect{x},t)}{\rho(\vect{x},t)}+\nu \Delta \vect{v}(\vect{x},t).
\end{align}
\end{subequations}

Taking the divergence of equation (\ref{analytical:navierStokesRewB}) we find
\begin{equation}
  \Delta p=\vect{\nabla}\cdot f(\vect{v}).
\end{equation}
But using the notation of (\ref{introduction:projectiondef}):
\begin{equation}
	\vect{\nabla}p=P(f(\vect{v})).
\end{equation}

The equations of Navier-Stokes are then equivalent to solving
\begin{equation}
  \partial_t \vect{v}(\vect{x} ,t)=f(\vect{v}(\vect{x},t))-P(f(\vect{v}(\vect{x},t))=(\eye-P)f(\vect{v}(\vect{x},t)).
\end{equation}
$(\eye-P)$ is a projector onto a divergence free space without changing the rotational.

\subsection{Mobile eulerian topology }
When topology is variable, we need to add another equation to tell us where the domain of computation lies.
A natural boundary condition is the free surface, which moves with the speed at a given point on the boundary.

We define $\lambda$ a system of representatives, and $\vect{\xi}_{\lambda}$ the position of points given by $\lambda$.
The domain is given by the union of $\vect{\xi}$ for all $\lambda$.

The Navier-Stokes equations are now
\begin{subequations}
\begin{align}
\label{analytical:navierStokesMobA}
\vect{\nabla} \cdot \vect{v}(\vect{x} ,t)&=0,\\
\label{analytical:navierStokesMobB}
\partial_t \vect{v}(\vect{x} ,t)+(\vect{v}(\vect{x},t)\cdot\vect{\nabla} ) \vect{v}(\vect{x} ,t)&=-\vect{\nabla} p(\vect{x},t)+\frac{\vect{F}(\vect{x},t)}{\rho(\vect{x},t)}+\nu \vect{\Delta} \vect{v}(\vect{x},t),\\
\partial_t \vect{\xi}_{\lambda}(t)&=\vect{v}(\vect{\xi}_{\lambda},t).
\end{align}
\end{subequations}
We can as in section \ref{analytical:fixe_eulerian} eliminate the constraint on the divergence with the same notation
\begin{subequations}
\begin{align}
\partial_t \vect{v}(\vect{x} ,t)&=f(\vect{v}(\vect{x},t))+P(f(\vect{v}(\vect{x},t))=(\eye+P)f(\vect{v}(\vect{x},t)),\\
\partial_t \vect{\xi}_{\lambda}(t)&=\vect{v}(\vect{\xi}_{\lambda},t)\label{ana:move:equ}.
\end{align}
\end{subequations}

\subsection{Lagrangian form of Navier-Stokes equations}
\label{analytical:convectif}
We will in this section rewrite (\ref{analytical:navierStokesB}) in another form.
We begin to define a new position system (which is similar to the particle position used in the previous section) $\vect{\xi}_{\lambda}(t)$:
\begin{subequations}
\begin{align}
 \partial_t \vect{\xi}_{\lambda}(t)&=\vect{v}(\vect{\xi}_{\lambda}(t),t),\\
 \vect{\xi}_{\lambda}(t_0)&=\vect{\xi}^{0}_{\lambda}.
\end{align}
\end{subequations}
Where $\lambda$ is a system of representatives that is initialized by an initial condition.
For a given $\lambda$, $\vect{\xi}_{\lambda}(t)$ follows the speed at the given point.
This position system is called particle coordinate or Lagrangian.
$\vect{\xi}_\lambda(t)$ is called the characteristic.

We then can define the Lagrangian speed
\begin{equation}
 \vect{u}_{\lambda}(t)=\vect{v}(\vect{\xi}_{\lambda}(t),t).
\end{equation}
The Lagrangian speed is the speed following a characteristic.

Taking the total derivative (the derivative of everything that depend on time) of the Lagrangian speed with respect of time, we have
\begin{equation}
\frac{d \vect{u}_{\lambda}(t)}{d t}=\frac{d \vect{v}(\xi_{\lambda},t)}{d t}=\partial_t \vect{v}+\left(\frac{\partial \vect{\xi}_{\lambda}(t)}{\partial t}\cdot\vect{\nabla}\right)\vect{v}.
\end{equation}
This can be rewritten as
\begin{equation}
\frac{d \vect{u}_{\lambda}(t)}{d t}=\partial_t \vect{v}+\left(\vect{v} \cdot\vect{\nabla}\right)\vect{v}.
\end{equation}
This is exactly the lhs of equation (\ref{analytical:navierStokesB}).

We then can rewrite the Navier-Stokes equations as
\begin{subequations}
\begin{align}
\label{analytical:navierStokesLagA}
\vect{\nabla} \cdot \vect{u}_{\lambda}(t)&=0,\\
\label{analytical:navierStokesLagB}
\frac{d \vect{u}_{\lambda}(t)}{d t}&=-\vect{\nabla} p(\vect{\xi}_{\lambda}(t),t)+\frac{\vect{F}(\vect{\xi}_{\lambda}(t),t)}{\rho(\vect{x},t)}+\nu \Delta \vect{u}_{\lambda}(t).
\end{align}
\end{subequations}

\begin{remark}
We remark that with this notation some expression like spatial derivative of $\vect{u}$ is not defined.
In an analytical procedure, we can transform $\vect{u}$ to $\vect{v}$ with the help of the expression of $\vect{\xi}$, then the derivative is well defined.
But in a numerical procedure, the point where the values are known are particles. This requires us to define derivatives on unstructured grids
or to interpolate at grid points from the particles.
\end{remark}

\section{Boundary conditions}

We will be interested in two kinds of boundary conditions.
\begin{description}
 \item[\emph{Inflow} Boundary:]
 An inflow boundary is a source of water.
 \item[\emph{Free surface} Boundary:]
 A free surface boundary is a freely moving boundary where the other fluid is inert (no density and no viscosity).
\end{description}

\subsection{Inflow}

Inflow boundary conditions have a constant prescribed speed.
This gives the following constraint on pressure:
\begin{equation}
\nabla p(\vect{x},t)=f(\vect{v}(\vect{x},t)).
\end{equation}

The following alternative boundary conditions have been used.
Set the speed on the surface artificially to constant and use homogeneous Neumann boundary conditions for pressure.


\subsection{Free surface}
\label{ana:free:surface}
To define the free surface boundary conditions, we need to define the stress tensor
\begin{equation}
	\sigma_{ij}=-p \delta_{ij}+\nu\left(\frac{\partial v_{i}}{\partial x_{j}}+\frac{\partial v_{j}}{\partial x_{i}}\right).
\end{equation}
The boundary conditions are then given by (for 3d):
\begin{align}
	\sum_{i,j}\sigma_{ij}n_{i}n_{j}&=0,\\
	\sum_{i,j}\sigma_{ij}t^{1}_{i}n_{j}&=0,\\
	\sum_{i,j}\sigma_{ij}t^{2}_{i}n_{j}&=0.\\
\end{align}
where $\vect{n}$ is the normal vector and $\vect{t}^1$ and $\vect{t}^2$ are two non-colinear tangent vectors.
For 2d only one tangent vector is needed.

The first equation can be used to find boundary conditions for $p$.
The other equations only depend on speed, because $\vect{n}$ is orthogonal to $\vect{t}^{1}$ and $\vect{t}^{2}$.
The value of $\nu$ is not relevant for the condition (a global non zero constant).
The constraint is linear.